outside yelloweye

### 2021-05-20 analysis/misc/yelloweye-south-outside.Rmd
exploring model structures for outside S HBLL

---
title: "outside halibut"
author: "Philina English"
date: "18/05/2021"
output: html_document
---

@@

  plot_map <- function(dat, column = "est") {
    ggplot(dat, aes_string("X", "Y", colour = column)) +
      geom_point(alpha = 0.5) +
      coord_fixed()+
      gfplot::theme_pbs() 
  }

@@

# combine event and hook data if not done already
```{r}
hook_adjusted_data <- "data-generated/halibut_hook_adjusted_36.rds"
if (!file.exists(hook_adjusted_data)) {
data <- readRDS("data/halibut-hook-dat36.rds")
d1 <- readRDS("data/halibut-surv-sets.rds") %>% filter(survey_abbrev=="HBLL OUT S")

@@

saveRDS(hook_adjusted_data, file = "data-generated/halibut_hook_adjusted_36.rds")
}
```

# add block ids and offsets
```{r}
f <- "data-generated/halibut-surv-36-blocks.rds"
if (file.exists(f)) {
  d <- readRDS(f)
} else {
d <- readRDS("data/halibut-surv-sets.rds") %>% filter(survey_abbrev=="HBLL OUT S")
  block_ids <- readRDS("data/survey-blocks-36.rds") 
  d <- dplyr::left_join(d, block_ids)
  d$block_designation <- as.numeric(d$block_designation)
  saveRDS(d, file = "data-generated/halibut-surv-36-blocks.rds")
}

hook <- readRDS("data-generated/halibut_hook_adjusted_36.rds") %>%
  select(
    survey_series_id, year, fishing_event_id, total_hooks, prop_bait_hooks,
    hook_adjust_factor, expected_catch
  )

@@

@@

# just depth
```{r}
# m0 <- sdmTMB(
#       formula = catch_count ~ 0 + as.factor(year) +
#           depth_scaled + I(depth_scaled^2),
#       data = d_utm,
#       spde = sp,
#       time = "year",
#       silent = FALSE,
#       anisotropy = TRUE,
#       #cores = cores,
#       reml = TRUE,
#       family = nbinom2(link = "log")
#     )
# saveRDS(m0, file = "models/halibut-m0-depth-200kn.rds")
m0  <-  readRDS(file = "models/halibut-m0-depth-200kn.rds")
m0
get_diag(m0)
```

```{r}
d_utm$offset <- d_utm$offset_area_hook

m1 <- sdmTMB(
      formula = catch_count ~ 0 + as.factor(year) + depth_scaled + I(depth_scaled^2) + offset,
      data = d_utm,
      spde = sp,
      time = "year",
      silent = FALSE,
      anisotropy = TRUE,
      #cores = cores,
      reml = TRUE,
      family = nbinom2(link = "log")
    )
saveRDS(m1, file = "models/halibut-m1-offset-area-hook-200kn.rds")

# m1  <- readRDS(file = "models/halibut-m1-offset-area-hook-200kn.rds")
m1
get_diag(m1)
```

# estimate effect of hook area 
```{r}
m1b <- sdmTMB(
      formula = catch_count ~ 0 + as.factor(year) + depth_scaled + I(depth_scaled^2) + offset_area_hook,
      data = d_utm,
      spde = sp,
      time = "year",
      silent = FALSE,
      anisotropy = TRUE,
      #cores = cores,
      reml = TRUE,
      family = nbinom2(link = "log")
    )
saveRDS(m1b, file = "models/halibut-m1b-area-hook-200kn.rds")
# m1b  <- readRDS(file = "models/halibut-m1b-area-hook-200kn.rds")
m1b
get_diag(m1b)
```

# area swept worse than area hook
```{r}
# d_utm$offset <- d_utm$offset_area_swept
# 
# m2 <- sdmTMB(
#       formula = catch_count ~ 0 + as.factor(year) +
#           depth_scaled + I(depth_scaled^2) + offset,
#       data = d_utm,
#       spde = sp,
#       time = "year",
#       silent = FALSE,
#       anisotropy = TRUE,
#       #cores = cores,
#       reml = TRUE,
#       family = nbinom2(link = "log")
#     )
# saveRDS(m2, file = "models/halibut-m2-offset-area-swept-200kn.rds")
m2  <-  readRDS(file = "models/halibut-m2-offset-area-swept-200kn.rds")
m2
get_diag(m2)
```

# estimate effect of area swept better AIC but worse MSE than area hook 
```{r}
m2b <- sdmTMB(
      formula = catch_count ~ 0 + as.factor(year) +
          depth_scaled + I(depth_scaled^2) + offset_area_swept,
      data = d_utm,
      spde = sp,
      time = "year",
      silent = FALSE,
      anisotropy = TRUE,
      #cores = cores,
      reml = TRUE,
      family = nbinom2(link = "log")
    )
saveRDS(m2b, file = "models/halibut-m2b-area-swept-200kn.rds")

# m2b  <-  readRDS(file = "models/halibut-m2b-area-swept-200kn.rds")
m2b
get_diag(m2b)
```
# just an estimate for the hook adjustment factor
## second best AIC to model predicting effect of area swept
```{r}
# m3 <- sdmTMB(
#       formula = catch_count ~ 0 + as.factor(year) +
#           depth_scaled + I(depth_scaled^2) +
#           hook_adjust_factor,
#       data = d_utm,
#       spde = sp,
#       time = "year",
#       silent = FALSE,
#       anisotropy = TRUE,
#       #cores = cores,
#       reml = TRUE,
#       family = nbinom2(link = "log")
#     )
# saveRDS(m3, file = "models/halibut-m3-hook-factor-200kn.rds")

m3  <- readRDS(file = "models/halibut-m3-hook-factor-200kn.rds")
m3
get_diag(m3)
```

# use swept area offset and estimate hook adjustment
```{r}
d_utm$offset <- d_utm$offset_area_swept

m4 <- sdmTMB(
      formula = catch_count ~ 0 + as.factor(year)+
          depth_scaled + I(depth_scaled^2) +
          hook_adjust_factor +
          offset,
      data = d_utm,
      spde = sp,
      time = "year",
      silent = FALSE,
      anisotropy = TRUE,
      #cores = cores,
      reml = TRUE,
      family = nbinom2(link = "log")
    )
saveRDS(m4, file = "models/halibut-m4-hookfact-offset-area-swept-200kn.rds")
# m4  <-  readRDS(file = "models/halibut-m4-hookfact-offset-area-swept-200kn.rds")
m4
get_diag(m4)
```

# estimate swept area and hook adjustment
```{r}
# m4b <- sdmTMB(
#       formula = catch_count ~ 0 + as.factor(year)+
#           depth_scaled + I(depth_scaled^2) +
#           hook_adjust_factor +
#           offset_area_swept,
#       data = d_utm,
#       spde = sp,
#       time = "year",
#       silent = FALSE,
#       anisotropy = TRUE,
#       #cores = cores,
#       reml = TRUE,
#       family = nbinom2(link = "log")
#     )
# saveRDS(m4b, file = "models/halibut-m4b-hookfact-area-swept-200kn.rds")
m4b  <-  readRDS(file = "models/halibut-m4b-hookfact-area-swept-200kn.rds")
m4b
get_diag(m4b)
```

@@

# estimate swept area and hook adjustment
```{r}
m4c <- sdmTMB(
      formula = catch_count ~ 0 + as.factor(year)+
          depth_scaled + I(depth_scaled^2) +
          hook_adjust_factor +
          offset_area_swept,
      data = d_utm,
      spde = sp4,
      time = "year",
      silent = FALSE,
      anisotropy = TRUE,
      #cores = cores,
      reml = TRUE,
      family = nbinom2(link = "log")
    )
saveRDS(m4c, file = "models/halibut-m4c-hookfact-area-swept-400kn.rds")
# m4c  <-  readRDS(file = "models/halibut-m4c-hookfact-area-swept-400kn.rds")
m4c
get_diag(m4c)
```

# offset of hook_count
```{r}
d_utm$offset <- log(d_utm$hook_count)

m5 <- sdmTMB(
      formula = catch_count ~ 0 + as.factor(year) + # hook_adjust_factor + # currently leaving out
          depth_scaled + I(depth_scaled^2) +
          offset,
      data = d_utm,
      spde = sp4,
      ar1_fields = T,
      time = "year",
      silent = FALSE,
      anisotropy = TRUE,
      #cores = cores,
      reml = TRUE,
      family = nbinom2(link = "log")
    )
saveRDS(m5, file = "models/halibut-m5-hook-count-400kn-AR1.rds")
# m5  <-  readRDS(file = "models/halibut-m5-hook-count-400kn-AR1.rds")
m5
tidy(m5, "ran_pars",conf.int = TRUE)
get_diag(m5)
```

# spatial only model
```{r}
# d_utm$offset <- log(d_utm$hook_count)
# 
# m5s <- sdmTMB(
#       formula = catch_count ~ 0 + as.factor(year) + # hook_adjust_factor + # currently leaving out
#           depth_scaled + I(depth_scaled^2) +
#           offset,
#       data = d_utm,
#       spde = sp4,
#       # ar1_fields = T,
#       # time = "year",
#       silent = FALSE,
#       anisotropy = TRUE,
#       #cores = cores,
#       reml = TRUE,
#       family = nbinom2(link = "log")
#     )
# saveRDS(m5s, file = "models/halibut-m5-hook-count-400kn-spatial.rds")
m5s  <-  readRDS(file = "models/halibut-m5-hook-count-400kn-spatial.rds")
m5s
tidy(m5s, "ran_pars",conf.int = TRUE)
get_diag(m5s)
```

# Predictions from best model
```{r}
# p <- predict(m5, newdata = s_grid_utm, se_fit = F, re_form = NULL, return_tmb_object = T)
# saveRDS(p, "models/halibut-m5-AR1-predictions.rds")
```

@@

# p <- readRDS("models/halibut-m4c-AR1-predictions.rds")
p <- readRDS("models/halibut-m5-AR1-predictions.rds")
g <- plot_map(p$data, exp(est)) +
  scale_fill_viridis_c(trans = ggsidekick::fourth_root_power_trans(), option = "D") +
  labs(
    fill = "Estimated\nadjusted\ncount",
    size = "Observed\nadjusted\ncount"
  ) +
  geom_point(
    data = d_utm, pch = 21, mapping = aes(
      x = X, y = Y,
      size = (catch_count / hook_count)*100
    ),
    inherit.aes = FALSE, colour = "grey20", alpha = 0.5
  ) +
  scale_size_area(max_size = 7)
g
```

@@

# Add court defined area shape file

@@

# Predictions for CDA only
```{r}
m5  <-  readRDS(file = "models/halibut-m5-hook-count-400kn-AR1.rds")
p_cda <- predict(m5, newdata = cda_grid, se_fit = F, re_form = NULL, return_tmb_object = T)
saveRDS(p_cda, "models/halibut-m5-AR1-predictions-cda.rds")

i_cda <- get_index(p_cda)
saveRDS(i_cda, "models/halibut-m5-AR1-index-cda.rds")

@@

# Predictions outside CDA but within 3cd
```{r}
# m5  <-  readRDS(file = "models/halibut-m5-hook-count-400kn-AR1.rds")
p_3cd <- predict(m5, newdata = noncda_grid, se_fit = F, re_form = NULL, return_tmb_object = T)
saveRDS(p_3cd, "models/halibut-m5-AR1-predictions-3cd.rds")

i_3cd <- get_index(p_3cd)
saveRDS(i_3cd, "models/halibut-m5-AR1-index-3cd.rds")

@@

# or without time
```{r}
m5s  <-  readRDS(file = "models/halibut-m5-hook-count-400kn-spatial.rds")
sp_cda <- predict(m5s, newdata = cda_grid, se_fit = F, re_form = NULL, return_tmb_object = T)
saveRDS(ps, "models/halibut-m5-spatial-predictions-cda.rds")

@@

sp_noncda <- predict(m5s, newdata = noncda_grid, se_fit = F, re_form = NULL, return_tmb_object = T)
saveRDS(ps, "models/halibut-m5-spatial-predictions-cda.rds")

i_noncda_sp <- get_index(sp_noncda)
# saveRDS(i_sp, "models/halibut-m5-cda-spatial-index.rds")

@@

# map all of S HBLL
```{r}
# creates coast lines for area defined in lat lon
load_coastll <- function(xlim_ll, ylim_ll, utm_zone, buffer = 0) {
  data("nepacLLhigh", package = "PBSmapping", envir = environment())
  np <- PBSmapping::clipPolys(nepacLLhigh,
    xlim = xlim_ll + c(-buffer, buffer),
    ylim = ylim_ll + c(-buffer, buffer)
  )
  # ll2utm(np, utm_zone = utm_zone)
}

@@

# p_hal <- readRDS("models/halibut-m4c-AR1-predictions.rds")
p_hal <- readRDS("models/halibut-m5-AR1-predictions.rds")
gh <- ggplot(data = focal_area) +
  # plot predictions from best model with mean swept area and hook adjustment factor
  geom_tile(data = filter(p_hal$data, year == 2020), 
    aes(longitude, latitude, fill = exp(est)), width = 0.06, height = 0.025) +
  scale_fill_viridis_c(trans = ggsidekick::fourth_root_power_trans(), 
    option = "D", limits=c(0, 40), na.value = "yellow") +
  labs(
    fill = "Estimated\nadjusted\ncount",
    size = "Observed\nadjusted\ncount"
  ) +
  geom_point(
    data = filter(d_utm, year == 2020), #just most recent year
    # data = d_utm, # plot all years together
    pch = 21, mapping = aes(
      longitude, latitude,
      size = (catch_count / hook_count)*100
    ),
    inherit.aes = FALSE, colour = "grey10", alpha = 0.4
  ) +  scale_size_area(max_size = 4) + 
  geom_line(# add major management region boundaries
    data = bound3CDnorth,
    aes(X, Y), colour = "black", lty = 2, 
    inherit.aes = F
  ) +
  annotate("text", x = -129.5, y = 50.25, label = "3CD") +
  annotate("text", x = -129.5, y = 51.25, label = "5AB") +
  geom_sf(colour = "red", fill = NA) + # add focal area
  geom_polygon( # add coast outline
      data = coast, aes(x = X, y = Y, group = PID),
      fill = "grey87", col = "grey70", lwd = 0.2
    ) +
  xlab("")+ ylab("")+ # remove axis labels
  coord_sf(expand = F, ylim=c(48.5, 52.1))  +# trim points beyond grid boundaries
  ggtitle("HBLL observed and predicted Pacific Halibut catch per 100 hooks")
gh
ggsave("halibut-m5-AR1-2020-predictions.png", width = 6, height = 6)
```

```{r}
gh3cd <- ggplot(data = focal_area) +
   geom_tile(data = filter(p_cda$data, year == 2020),
   aes(longitude, latitude, fill = exp(est)), width = 0.06, height = 0.025) +
   geom_tile(data = filter(p_3cd$data, year == 2020),
    aes(longitude, latitude, fill = exp(est)), width = 0.05, height = 0.025) +
  scale_fill_viridis_c(
    trans = ggsidekick::fourth_root_power_trans(),
    limits=c(0, 7), na.value = "yellow",
    option = "D") +
  labs(
    fill = "Estimated\ncount\nin 3CD",
    size = "Observed\ncount"
  ) +
  geom_point(# plot all years together
    data = d_utm, pch = 21, mapping = aes(
      longitude, latitude,
      size = (catch_count / hook_count)*100
    ),
    inherit.aes = FALSE, colour = "grey10", alpha = 0.4
  ) +  scale_size_area(max_size = 4) + 
  geom_sf(colour = "red", fill = NA) + # add focal area
  geom_polygon( # add coast outline
      data = coast, aes(x = X, y = Y, group = PID),
      fill = "grey87", col = "grey70", lwd = 0.2
    ) +
  xlab("")+ ylab("")+ # remove axis labels
  coord_sf(expand = F, ylim=c(48.4, 50.6), xlim=c(-128.5, -125.15))  +# trim points beyond grid boundaries
  ggtitle("HBLL observed and predicted Pacific Halibut catch per 100 hooks")
gh3cd
ggsave("halibut-m5-AR1-2020-predictions-3cd.png", width = 6, height = 6)
```

@@

gratio <- ggplot(data = focal_area) +
  # plot predictions from best model with mean swept area and hook adjustment factor
  geom_tile(data = filter(ratio_df, year == 2020), 
    aes(longitude, latitude, fill = ye_per_hal), width = 0.06, height = 0.025) +
  scale_fill_viridis_c(trans = ggsidekick::fourth_root_power_trans(), 
    # limits=c(0, 40), na.value = "yellow", 
    option = "D"
    ) +
  labs(
    fill = "Yelloweye\ncaught\nper halibut",
    size = "Observed\nhalibut\ncatch"
  ) +
  geom_point(# plot all years together
    data = d_utm, pch = 21, mapping = aes(
      longitude, latitude,
      size = (catch_count / hook_count)*100
    ),
    inherit.aes = FALSE, colour = "grey10", alpha = 0.4
  ) +  scale_size_area(max_size = 4) + 
  geom_line(# add major management region boundaries
    data = bound3CDnorth,
    aes(X, Y), colour = "black", lty = 2, 
    inherit.aes = F
  ) +
  annotate("text", x = -129.5, y = 50.25, label = "3CD") +
  annotate("text", x = -129.5, y = 51.25, label = "5AB") +
  geom_sf(colour = "red", fill = NA) + # add focal area
  geom_polygon( # add coast outline
      data = coast, aes(x = X, y = Y, group = PID),
      fill = "grey87", col = "grey70", lwd = 0.2
    ) +
  xlab("")+ ylab("")+ # remove axis labels
  coord_sf(expand = F, ylim=c(48.5, 52.1))  +# trim points beyond grid boundaries
  ggtitle("Estimated number of Yelloweye Rockfish caught per halibut")
gratio
ggsave("ye-per-halibut-w-hal-catch.png", width = 6, height = 6)

>>>>

---
title: "outside yelloweye"
author: "Philina English"
date: "18/05/2021"
output: html_document
---

@@

  plot_map <- function(dat, column = "est") {
    ggplot(dat, aes_string("X", "Y", colour = column)) +
      geom_point(alpha = 0.25) +
      coord_fixed()+
      gfplot::theme_pbs() 
  }

@@

# combine event and hook data if not done already
```{r}
hook_adjusted_data <- "data-generated/yelloweye_hook_adjusted_36.rds"
if (!file.exists(hook_adjusted_data)) {
data <- readRDS("data/yelloweye-rockfish-hook-dat36.rds")
d1 <- readRDS("data/yelloweye-rockfish-surv-sets.rds") %>% filter(survey_abbrev=="HBLL OUT S")

@@

saveRDS(hook_adjusted_data, file = "data-generated/yelloweye_hook_adjusted_36.rds")
}
```

# add block ids and offsets
```{r}
f <- "data-generated/yelloweye-rockfish-surv-36-blocks.rds"
if (file.exists(f)) {
  d <- readRDS(f)
} else {
d <- readRDS("data/yelloweye-rockfish-surv-sets.rds") %>% filter(survey_abbrev=="HBLL OUT S")
  block_ids <- readRDS("data/survey-blocks-36.rds") 
  d <- dplyr::left_join(d, block_ids)
  d$block_designation <- as.numeric(d$block_designation)
  saveRDS(d, file = "data-generated/yelloweye-rockfish-surv-36-blocks.rds")
}

hook <- readRDS("data-generated/yelloweye_hook_adjusted_36.rds") %>%
  select(
    survey_series_id, year, fishing_event_id, total_hooks, prop_bait_hooks,
    hook_adjust_factor, expected_catch
  )

@@

# hb <- readRDS("data/halibut-surv-sets.rds") %>% filter(survey_abbrev=="HBLL OUT S")

@@

# just depth
```{r}
# m0 <- sdmTMB(
#       formula = catch_count ~ 0 + as.factor(year) + 
#           depth_scaled + I(depth_scaled^2),
#       data = d_utm,
#       spde = sp,
#       time = "year",
#       silent = FALSE,
#       anisotropy = TRUE,
#       #cores = cores,
#       reml = TRUE,
#       family = nbinom2(link = "log")
#     )
# saveRDS(m0, file = "models/ye-m0-depth-200kn.rds")
m0  <-  readRDS(file = "models/ye-m0-depth-200kn.rds")
m0

get_diag(m0)
```

```{r}
# d_utm$offset <- d_utm$offset_area_hook
# 
# m1 <- sdmTMB(
#       formula = catch_count ~ 0 + as.factor(year) + depth_scaled + I(depth_scaled^2) + offset,
#       data = d_utm,
#       spde = sp,
#       time = "year",
#       silent = FALSE,
#       anisotropy = TRUE,
#       #cores = cores,
#       reml = TRUE,
#       family = nbinom2(link = "log")
#     )
# saveRDS(m1, file = "models/ye-m1-offset-area-hook-200kn.rds")

m1  <- readRDS(file = "models/ye-m1-offset-area-hook-200kn.rds")
m1
get_diag(m1)
```

# estimate effect of hook area 
```{r}
# m1b <- sdmTMB(
#       formula = catch_count ~ 0 + as.factor(year) + depth_scaled + I(depth_scaled^2) + offset_area_hook,
#       data = d_utm,
#       spde = sp,
#       time = "year",
#       silent = FALSE,
#       anisotropy = TRUE,
#       #cores = cores,
#       reml = TRUE,
#       family = nbinom2(link = "log")
#     )
# saveRDS(m1b, file = "models/ye-m1b-area-hook-200kn.rds")
m1b  <- readRDS(file = "models/ye-m1b-area-hook-200kn.rds")
m1b
get_diag(m1b)
```

```{r}
# d_utm$offset <- d_utm$offset_area_swept

# m2 <- sdmTMB(
#       formula = catch_count ~ 0 + as.factor(year) + 
#           depth_scaled + I(depth_scaled^2) + offset,
#       data = d_utm,
#       spde = sp,
#       time = "year",
#       silent = FALSE,
#       anisotropy = TRUE,
#       #cores = cores,
#       reml = TRUE,
#       family = nbinom2(link = "log")
#     )
# saveRDS(m2, file = "models/ye-m2-offset-area-swept-200kn.rds")
m2  <-  readRDS(file = "models/ye-m2-offset-area-swept-200kn.rds")
m2
get_diag(m2)
```
# estimate effect of area swept 
## best so far
```{r}
# m2b <- sdmTMB(
#       formula = catch_count ~ 0 + as.factor(year) + 
#           depth_scaled + I(depth_scaled^2) + offset_area_swept,
#       data = d_utm,
#       spde = sp,
#       time = "year",
#       silent = FALSE,
#       anisotropy = TRUE,
#       #cores = cores,
#       reml = TRUE,
#       family = nbinom2(link = "log")
#     )
# saveRDS(m2b, file = "models/ye-m2b-area-swept-200kn.rds")

m2b  <-  readRDS(file = "models/ye-m2b-area-swept-200kn.rds")
m2b
get_diag(m2b)
```
# just an estimate for the hook adjustment factor
## second best AIC to model predicting effect of area swept
```{r}
# m3 <- sdmTMB(
#       formula = catch_count ~ 0 + as.factor(year) + 
#           depth_scaled + I(depth_scaled^2) + 
#           hook_adjust_factor,
#       data = d_utm,
#       spde = sp,
#       time = "year",
#       silent = FALSE,
#       anisotropy = TRUE,
#       #cores = cores,
#       reml = TRUE,
#       family = nbinom2(link = "log")
#     )
# saveRDS(m3, file = "models/ye-m3-hook-factor-200kn.rds")

m3  <- readRDS(file = "models/ye-m3-hook-factor-200kn.rds")
m3
get_diag(m3)
```

# use swept area offset and estimate hook adjustment
```{r}
# d_utm$offset <- d_utm$offset_area_swept
# 
# m4 <- sdmTMB(
#       formula = catch_count ~ 0 + as.factor(year)+
#           depth_scaled + I(depth_scaled^2) + 
#           hook_adjust_factor +
#           offset,
#       data = d_utm,
#       spde = sp,
#       time = "year",
#       silent = FALSE,
#       anisotropy = TRUE,
#       #cores = cores,
#       reml = TRUE,
#       family = nbinom2(link = "log")
#     )
# saveRDS(m4, file = "models/ye-m4-hookfact-offset-area-swept-200kn.rds")
m4  <-  readRDS(file = "models/ye-m4-hookfact-offset-area-swept-200kn.rds")
m4
get_diag(m4)
```

# estimate swept area and hook adjustment
```{r}
# m4b <- sdmTMB(
#       formula = catch_count ~ 0 + as.factor(year)+
#           depth_scaled + I(depth_scaled^2) +
#           hook_adjust_factor +
#           offset_area_swept,
#       data = d_utm,
#       spde = sp,
#       time = "year",
#       silent = FALSE,
#       anisotropy = TRUE,
#       #cores = cores,
#       reml = TRUE,
#       family = nbinom2(link = "log")
#     )
# saveRDS(m4b, file = "models/ye-m4b-hookfact-area-swept-200kn.rds")
m4b  <-  readRDS(file = "models/ye-m4b-hookfact-area-swept-200kn.rds")
m4b
get_diag(m4b)
```

@@

# estimate swept area and hook adjustment
```{r}
m4c <- sdmTMB(
      formula = catch_count ~ 0 + as.factor(year)+
          depth_scaled + I(depth_scaled^2) +
          hook_adjust_factor +
          offset_area_swept,
      data = d_utm,
      spde = sp4,
      ar1_fields = T,
      time = "year",
      silent = FALSE,
      anisotropy = TRUE,
      #cores = cores,
      reml = TRUE,
      family = nbinom2(link = "log")
    )
saveRDS(m4c, file = "models/ye-m4c-hookfact-area-swept-400kn-AR1.rds")
m4c  <-  readRDS(file = "models/ye-m4c-hookfact-area-swept-400kn-AR1.rds")
m4c
tidy(m4c, "ran_pars",conf.int = TRUE)
get_diag(m4c)
```

# offset of hook_count
```{r}
# d_utm$offset <- log(d_utm$hook_count)
# 
# m5 <- sdmTMB(
#       formula = catch_count ~ 0 + as.factor(year) + # hook_adjust_factor + # currently leaving out
#           depth_scaled + I(depth_scaled^2) + 
#           offset,
#       data = d_utm,
#       spde = sp4,
#       ar1_fields = T,
#       time = "year",
#       silent = FALSE,
#       anisotropy = TRUE,
#       #cores = cores,
#       reml = TRUE,
#       family = nbinom2(link = "log")
#     )
saveRDS(m5, file = "models/ye-m5-hook-count-400kn-AR1.rds")
# m5  <-  readRDS(file = "models/ye-m5-hook-count-400kn-AR1.rds")
m5
tidy(m5, "ran_pars",conf.int = TRUE)
get_diag(m5)
```

# spatial only model
```{r}
d_utm$offset <- log(d_utm$hook_count)

m5s <- sdmTMB(
      formula = catch_count ~ 0 + as.factor(year) + # hook_adjust_factor + # currently leaving out
          depth_scaled + I(depth_scaled^2) +
          offset,
      data = d_utm,
      spde = sp4,
      # ar1_fields = T,
      # time = "year",
      silent = FALSE,
      anisotropy = TRUE,
      #cores = cores,
      reml = TRUE,
      family = nbinom2(link = "log")
    )
saveRDS(m5s, file = "models/ye-m5-hook-count-400kn-spatial.rds")
# m5s  <-  readRDS(file = "models/ye-m5-hook-count-400kn-spatial.rds")
m5s
tidy(m5s, "ran_pars",conf.int = TRUE)
get_diag(m5s)
```

# Predictions from best model
```{r}

# m5  <-  readRDS(file = "models/ye-m5-hook-count-400kn-AR1.rds")
# p <- predict(m5, newdata = s_grid_utm, se_fit = F, re_form = NULL, return_tmb_object = T)
# saveRDS(p, "models/ye-m5-AR1-predictions.rds")
```

@@

# p <- readRDS("models/ye-m4c-AR1-predictions.rds")
p <- readRDS("models/ye-m5-AR1-predictions.rds")
g <- plot_map(p$data, exp(est)) +
  scale_fill_viridis_c(trans = ggsidekick::fourth_root_power_trans(), option = "D") +
  labs(
    fill = "Estimated\nadjusted\ncount",
    size = "Observed\nadjusted\ncount"
  ) +
  geom_point(
    data = d_utm, pch = 21, mapping = aes(
      x = X, y = Y,
      size = (catch_count / hook_count)*100
    ),
    inherit.aes = FALSE, colour = "grey20", alpha = 0.5
  ) +
  scale_size_area(max_size = 7)
g
```

@@

# try adding shape files

@@

# Predictions for CDA only
```{r}
m5  <-  readRDS(file = "models/ye-m5-hook-count-400kn-AR1.rds")
p_cda <- predict(m5, newdata = cda_grid, se_fit = F, re_form = NULL, return_tmb_object = T)
# saveRDS(p_cda, "models/ye-m5-AR1-predictions-cda.rds")

i_cda <- get_index(p_cda)
# saveRDS(i_cda, "models/ye-m5-AR1-index-cda.rds")

@@

# Predictions outside CDA but within 3cd
```{r}
# m5  <-  readRDS(file = "models/ye-m5-hook-count-400kn-AR1.rds")
p_3cd <- predict(m5, newdata = noncda_grid, se_fit = F, re_form = NULL, return_tmb_object = T)
saveRDS(p_3cd, "models/ye-m5-AR1-predictions-3cd.rds")

i_3cd <- get_index(p_3cd)
saveRDS(i_3cd, "models/ye-m5-AR1-index-3cd.rds")

@@

# or without time
```{r}
m5s  <-  readRDS(file = "models/ye-m5-hook-count-400kn-spatial.rds")
sp_cda <- predict(m5s, newdata = cda_grid, se_fit = F, re_form = NULL, return_tmb_object = T)
saveRDS(ps, "models/ye-m5-spatial-predictions-cda.rds")

@@

sp_noncda <- predict(m5s, newdata = noncda_grid, se_fit = F, re_form = NULL, return_tmb_object = T)
saveRDS(ps, "models/ye-m5-spatial-predictions-cda.rds")

i_noncda_sp <- get_index(sp_noncda)
# saveRDS(i_sp, "models/ye-m5-cda-spatial-index.rds")

@@

# creates coast lines for area defined in lat lon
load_coastll <- function(xlim_ll, ylim_ll, utm_zone, buffer = 0) {
  data("nepacLLhigh", package = "PBSmapping", envir = environment())
  np <- PBSmapping::clipPolys(nepacLLhigh,
    xlim = xlim_ll + c(-buffer, buffer),
    ylim = ylim_ll + c(-buffer, buffer)
  )
  # ll2utm(np, utm_zone = utm_zone)
}

@@

# p <- readRDS("models/ye-m4c-AR1-predictions.rds")
p <- readRDS("models/ye-m5-AR1-predictions.rds")
gye <- ggplot(data = focal_area) +
  # plot predictions from best model with mean swept area and hook adjustment factor
  geom_tile(data = filter(p$data, year == 2020),
    aes(longitude, latitude, fill = exp(est)), width = 0.06, height = 0.025) +
  scale_fill_viridis_c(trans = ggsidekick::fourth_root_power_trans(), 
    option = "D", limits=c(0, 40), na.value = "yellow") +
  labs(
    fill = "Estimated\nadjusted\ncount",
    size = "Observed\nadjusted\ncount"
  ) +
  geom_point(# plot all years together
    data = d_utm, pch = 21, mapping = aes(
      longitude, latitude,
      size = (catch_count / hook_count)*100
    ),
    inherit.aes = FALSE, colour = "grey10", alpha = 0.4
  ) +  scale_size_area(max_size = 4) + 
  geom_line(# add major management region boundaries
    data = bound3CDnorth,
    aes(X, Y), colour = "black", lty = 2, 
    inherit.aes = F
  ) +
  annotate("text", x = -129.5, y = 50.25, label = "3CD") +
  annotate("text", x = -129.5, y = 51.25, label = "5AB") +
  geom_sf(colour = "red", fill = NA) + # add focal area
  geom_polygon( # add coast outline
      data = coast, aes(x = X, y = Y, group = PID),
      fill = "grey87", col = "grey70", lwd = 0.2
    ) +
  xlab("")+ ylab("")+ # remove axis labels
  coord_sf(expand = F, ylim=c(48.5, 52.1))  +# trim points beyond grid boundaries
  ggtitle("HBLL observed and predicted Yelloweye Rockfish catch per 100 hooks")
gye
# ggsave("ye-m5-AR1-2020-predictions.png", width = 6, height = 6)
```

```{r}
g3cd <- ggplot(data = focal_area) +
   geom_tile(data = filter(p_cda$data, year == 2020),
   aes(longitude, latitude, fill = exp(est)), width = 0.06, height = 0.025) +
   geom_tile(data = filter(p_3cd$data, year == 2020),
    aes(longitude, latitude, fill = exp(est)), width = 0.05, height = 0.025) +
  scale_fill_viridis_c(
    trans = ggsidekick::fourth_root_power_trans(),
    # limits=c(0, 40), na.value = "yellow",
    option = "D") +
  labs(
    fill = "Estimated\ncount\nin 3CD",
    size = "Observed\ncount"
  ) +
  geom_point(# plot all years together
    data = d_utm, pch = 21, mapping = aes(
      longitude, latitude,
      size = (catch_count / hook_count)*100
    ),
    inherit.aes = FALSE, colour = "grey10", alpha = 0.4
  ) +  scale_size_area(max_size = 4) + 
  # geom_line(# add major management region boundaries
  #   data = bound3CDnorth,
  #   aes(X, Y), colour = "black", lty = 2, 
  #   inherit.aes = F
  # ) +
  # annotate("text", x = -128, y = 49.5, label = "3CD") +
  geom_sf(colour = "red", fill = NA) + # add focal area
  geom_polygon( # add coast outline
      data = coast, aes(x = X, y = Y, group = PID),
      fill = "grey87", col = "grey70", lwd = 0.2
    ) +
  xlab("")+ ylab("")+ # remove axis labels
  coord_sf(expand = F, ylim=c(48.4, 50.6), xlim=c(-128.5, -125.15))  +# trim points beyond grid boundaries
  ggtitle("HBLL observed and predicted Yelloweye Rockfish catch per 100 hooks")
g3cd
ggsave("ye-m5-AR1-2020-predictions-3cd.png", width = 6, height = 6)
```

@@

gratio <- ggplot(data = focal_area) +
  # plot predictions from best model with mean swept area and hook adjustment factor
  geom_tile(data = filter(ratio_df, year == 2020), 
    aes(longitude, latitude, fill = ye_per_hal), width = 0.06, height = 0.025) +
  scale_fill_viridis_c(trans = ggsidekick::fourth_root_power_trans(), 
    # limits=c(0, 40), na.value = "yellow", 
    option = "D"
    ) +
  labs(
    fill = "Yelloweye\ncaught\nper halibut",
    size = "Observed\nyelloweye\ncatch"
  ) +
  geom_point(# plot all years together
    data = d_utm, pch = 21, mapping = aes(
      longitude, latitude,
      size = (catch_count / hook_count)*100
    ),
    inherit.aes = FALSE, colour = "grey10", alpha = 0.4
  ) +  scale_size_area(max_size = 4) + 
  geom_line(# add major management region boundaries
    data = bound3CDnorth,
    aes(X, Y), colour = "black", lty = 2, 
    inherit.aes = F
  ) +
  annotate("text", x = -129.5, y = 50.25, label = "3CD") +
  annotate("text", x = -129.5, y = 51.25, label = "5AB") +
  geom_sf(colour = "red", fill = NA) + # add focal area
  geom_polygon( # add coast outline
      data = coast, aes(x = X, y = Y, group = PID),
      fill = "grey87", col = "grey70", lwd = 0.2
    ) +
  xlab("")+ ylab("")+ # remove axis labels
  coord_sf(expand = F, ylim=c(48.5, 52.1))  +# trim points beyond grid boundaries
  ggtitle("Estimated number of Yelloweye Rockfish caught per halibut")
gratio
ggsave("ye-per-halibut.png", width = 6, height = 6)
